---
title: "03_Real_Data_Analysis"
author: "Yufang Wang"
date: "2025-03-18"
output: pdf_document
---

# Step 1: Set up the environment
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
setwd("C:/Users/Alisa_Wang/Desktop/MasterThesis/Code/")#C:/Users/Alisa_Wang/Desktop/MasterThesis/Code/")#G:/MasterThesis/Code/"
source("./function_file/Functions.R")
source("./function_file/stan_constructor.R")

install_and_load <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE)) {
      install.packages(pkg, dependencies = TRUE)
    }
    library(pkg, character.only = TRUE) }
  
}

# List of required packages
required_packages <- c("rstan", "MASS", "rstantools", "coda", "readr", "survival", "splines2", "dplyr", "simsurv", "Hmisc", "caret", "corrplot","ggplot2", "loo", "survival", "caret")
# Install and/or load packages
install_and_load(required_packages)
```


# Step 2: Organize and strafied the real-world data
```{r, echo=TRUE}
real_df <- readRDS("./Data/imputed_NOTR_DGF.rds")
# sum(!complete.cases(real_df))
# mean(real_df$status == "graftloss")
# mean(real_df$status == "alive")
# mean(real_df$status == "death")

# # Reversed KM for follow-up analysis
# fit <- survfit(Surv(real_df$time, real_df$status != "graftloss") ~ 1)
# plot(fit)
# abline(h = 0.5, col = "red", lty = 2)
# 
# str(real_df)
real_df <- real_df[, !names(real_df) %in% "ID"]

# set the factor variable being sum-to-zero encoded
contrasts(real_df$Recipientsex) <- contr.sum(levels(real_df$Recipientsex))
contrasts(real_df$Donorsex) <- contr.sum(levels(real_df$Donorsex))
contrasts(real_df$Smoking) <- contr.sum(levels(real_df$Smoking))
contrasts(real_df$InitialPrimaryDiseaseET_regroup) <- contr.sum(levels(real_df$InitialPrimaryDiseaseET_regroup))
contrasts(real_df$Donorcauseofdeath_group) <- contr.sum(levels(real_df$Donorcauseofdeath_group))


# center the numeric variables
real_df$Recipientage <- scale(real_df$Recipientage)
real_df$Dialysisdaysrenine <- scale(real_df$Dialysisdaysrenine)
real_df$Donorage <- scale(real_df$Donorage)
real_df$Lastcreatininemgdl <- scale(real_df$Lastcreatininemgdl)
real_df$Warm_ischaemic_period_1 <- scale(real_df$Warm_ischaemic_period_1)
real_df$Cold_ischaemic_period <- scale(real_df$Cold_ischaemic_period)
real_df$HLA_mismatch <- scale(real_df$HLA_mismatch)
real_df$DonorBMI <- scale(real_df$DonorBMI)
real_df$Recipient_BMI <- scale(real_df$Recipient_BMI)
#

# print(colnames(real_df)[factorName])

real_df <- real_df[, !(names(real_df) %in% "InitialOnmachineindicator")]
real_df$Donorcauseofdeath_group <- ifelse(real_df$Donorcauseofdeath_group == "Cardiac", "Cardiac", "Others")
real_df$Donorcauseofdeath_group <- factor(real_df$Donorcauseofdeath_group)
contrasts(real_df$Donorcauseofdeath_group) <- contr.sum(levels(real_df$Donorcauseofdeath_group))

real_df$InitialPrimaryDiseaseET_regroup <- ifelse(real_df$InitialPrimaryDiseaseET_regroup == "Diabetes Mellitus", "Diabetes", "Others")
real_df$InitialPrimaryDiseaseET_regroup <- factor(real_df$InitialPrimaryDiseaseET_regroup)
contrasts(real_df$InitialPrimaryDiseaseET_regroup) <- contr.sum(levels(real_df$InitialPrimaryDiseaseET_regroup))


combined_Data <- subset(real_df, select = -Warm_ischaemic_period_1)

combined_Data$noise1 <- rnorm(nrow(combined_Data), 0, 1)
combined_Data$noise2 <- rnorm(nrow(combined_Data), 1, 2)
combined_Data$time <- combined_Data$time + 0.05
```

# Descriptive data analysis
```{r, echo=TRUE}
library(dplyr)
combined_Data_encoded <- subset(combined_Data, select = -c(time, status)) %>%
  mutate(across(where(is.factor), as.character)) %>%
  mutate(across(where(is.character), ~ as.factor(.))) %>%
  model.matrix(~ . - 1, data = .) %>%
  as.data.frame() %>%
  mutate(across(everything(), ~ ifelse(. == 0, -1, .)))

# Step 3: Calculate correlation matrix
cor_matrix <- cor(combined_Data_encoded)

# Step 4: View the correlation matrix
print(cor_matrix)

# Optional: Visualize with a heatmap
library(corrplot)
corrplot(cor_matrix, method = "color", tl.cex = 0.7, type = "lower")
```



```{r, echo=TRUE}
combined_Data_Simple <- subset(combined_Data, select = -c(Smoking, Donorcauseofdeath_group, Dialysisdaysrenine, Recipientsex, Recipient_BMI))
# set the factor variable being sum-to-zero encoded
# contrasts(real_df_simple$Donorsex) <- contr.sum(levels(real_df_simple$Donorsex))
# contrasts(real_df_simple$TypecadavericDBDDCD) <- contr.sum(levels(real_df_simple$TypecadavericDBDDCD))
# contrasts(real_df_simple$InitialPrimaryDiseaseET_regroup) <- contr.sum(levels(real_df_simple$InitialPrimaryDiseaseET_regroup))
# 
# 
# 
# # center the numeric variables
# real_df_simple$Recipientage <- scale(real_df_simple$Recipientage)
# real_df_simple$Donorage <- scale(real_df_simple$Donorage)
# real_df_simple$Lastcreatininemgdl <- scale(real_df_simple$Lastcreatininemgdl)
# real_df_simple$Cold_ischaemic_period <- scale(real_df_simple$Cold_ischaemic_period)
# real_df_simple$HLA_mismatch <- scale(real_df_simple$HLA_mismatch)
# real_df_simple$DonorBMI <- scale(real_df_simple$DonorBMI)

# 
# combined_Data_Simple <- real_df_simple
```

# Check whether or not each level have sufficient events
```{r, echo=TRUE}
# Identify categorical variables
categorical_vars <- names(combined_Data)[sapply(combined_Data, function(x) is.factor(x) || is.character(x))]

# Print the categorical variables
print(categorical_vars)

for (cvars in categorical_vars){
  if (cvars != "status"){
    print(cvars)
    print(table(combined_Data$status == "graftloss", combined_Data[[cvars]]))
  }
}
```

# Section 1: include all variables and see whether or not it is possible for the get the model converenge

## Step 4: Verify the existence of interaction effects
```{r, echo=TRUE}
combined_Data_int <- includingInter(combined_Data)
# combined_Data_int <- subset(combined_Data_int, select = -c(id))

combined_Data_Simple_int <- includingInter(combined_Data_Simple)
#combined_Data_Simple_int <- subset(combined_Data_Simple_int, select = -c(id))
```
# Data analysis
```{r, echo=TRUE}
Data_bSplines <- stan_data_Constructer_BSpline(
  training_dataset = combined_Data_int,
  testing_dataset = NULL,
  obs_window = 3416,
)
bSplines_model <-
  Bayesian_Survival_model(
    stan_data = Data_bSplines,
    baseline_modelling = "bSplines",
    shrinkage = TRUE,
    withPrediction = FALSE,
    niter = 10000,
    nwarmup = 2000,
    thin = 10,
    chains = 1
  )
    
# Extract the model result
bSplines_model_results = Bayesian_Survival_result_Extract(
  bayesian_model_fit = bSplines_model,
  model_type = "bSplines",
  criteria = c(
    "DesignCoefficients",
    "variableSelection",
    "baseline"
  ),
  uncertainty = TRUE
)
    
bSplines_model_beta <- bSplines_model_results$Beta_bayesian_est[, "mean"]
bSplines_model_variableSelection <- bSplines_model_results$variableSelection
bSplines_model_coefUB <- bSplines_model_results$Beta_bayesian_estUB
bSplines_model_coefLB <- bSplines_model_results$Beta_bayesian_estLB

 #-----------MODEL 1.3: PL model----------
Data_PL <- stan_data_Constructer_PL(
  training_dataset = combined_Data_int,
  testing_dataset = NULL,
  havingInt = TRUE
)


PL_model <- Bayesian_Survival_model(
  stan_data = Data_PL,
  baseline_modelling = "none",
  shrinkage = TRUE,
  withPrediction = FALSE,
  niter = 10000,
  nwarmup = 2000,
  thin = 10,
  chains = 1
)

PL_results = Bayesian_Survival_result_Extract(
  bayesian_model_fit = PL_model,
  model_type = "none",
  criteria = c("DesignCoefficients",
               "variableSelection"),
  uncertainty = TRUE
)
PL_beta <- PL_results$Beta_bayesian_est[, "mean"]
PL_variableSelection <- PL_results$variableSelection
PL_coefUB <- PL_results$Beta_bayesian_estUB
PL_coefLB <- PL_results$Beta_bayesian_estLB


#----------NO SHRINKAGE---------------------
bSplines_model_noshrinkage <-
  Bayesian_Survival_model(
    stan_data = Data_bSplines,
    baseline_modelling = "bSplines",
    shrinkage = FALSE,
    withPrediction = FALSE,
    havingInt = TRUE,
    niter = 10000,
    nwarmup = 2000,
    thin = 10,
    chains = 1
  )
    
# Extract the model result
bSplines_model_noshrinkage_results = Bayesian_Survival_result_Extract(
  bayesian_model_fit = bSplines_model_noshrinkage,
  model_type = "bSplines",
  criteria = c(
    "DesignCoefficients",
    "variableSelection",
    "baseline"
  ),
  uncertainty = TRUE
)
    
bSplines_model_noshrinkage_beta <- bSplines_model_noshrinkage_results$Beta_bayesian_est[, "mean"]
bSplines_model_noshrinkage_variableSelection <- bSplines_model_noshrinkage_results$variableSelection
bSplines_model_noshrinkage_coefUB <- bSplines_model_noshrinkage_results$Beta_bayesian_estUB
bSplines_model_noshrinkage_coefLB <- bSplines_model_noshrinkage_results$Beta_bayesian_estLB


PL_model_noshrinkage <- Bayesian_Survival_model(
  stan_data = Data_PL,
  baseline_modelling = "none",
  shrinkage = FALSE,
  withPrediction = FALSE,
  havingInt = TRUE,
  niter = 10000,
  nwarmup = 2000,
  thin = 10,
  chains = 1
)

PL_noshrinkage_results = Bayesian_Survival_result_Extract(
  bayesian_model_fit = PL_model_noshrinkage,
  model_type = "none",
  criteria = c("DesignCoefficients",
               "variableSelection"),
  uncertainty = TRUE
)
PL_noshrinkage_beta <- PL_noshrinkage_results$Beta_bayesian_est[, "mean"]
PL_noshrinkage_variableSelection <- PL_noshrinkage_results$variableSelection
PL_noshrinkage_coefUB <- PL_noshrinkage_results$Beta_bayesian_estUB
PL_noshrinkage_coefLB <- PL_noshrinkage_results$Beta_bayesian_estLB
```
#################FOR THE SIMPLIFIED MODEL############

```{r, echo=TRUE}
Data_bSplines_Simple <- stan_data_Constructer_BSpline(
  training_dataset = combined_Data_Simple_int,
  testing_dataset = NULL,
  obs_window = 3415,
  log_lik_saving = FALSE,
  havingInt = TRUE
)
bSplines_model_Simple <-
  Bayesian_Survival_model(
    stan_data = Data_bSplines_Simple,
    baseline_modelling = "bSplines",
    shrinkage = TRUE,
    withPrediction = FALSE,
    havingInt = TRUE,
    niter = 10000,
    nwarmup = 2000,
    thin = 10,
    chains = 1
  )
    
# Extract the model result
bSplines_model_results_Simple = Bayesian_Survival_result_Extract(
  bayesian_model_fit = bSplines_model_Simple,
  model_type = "bSplines",
  criteria = c(
    "DesignCoefficients",
    "variableSelection",
    "baseline"
  ),
  uncertainty = TRUE
)
    
bSplines_model_Simple_beta <- bSplines_model_results_Simple$Beta_bayesian_est[, "mean"]
bSplines_model_Simple_variableSelection <- bSplines_model_results_Simple$variableSelection
bSplines_model_Simple_coefUB <- bSplines_model_results_Simple$Beta_bayesian_estUB
bSplines_model_Simple_coefLB <- bSplines_model_results_Simple$Beta_bayesian_estLB

 #-----------MODEL 1.3: PL model----------
Data_PL_Simple <- stan_data_Constructer_PL(
  training_dataset = combined_Data_Simple_int,
  testing_dataset = NULL,
  havingInt = TRUE
)


PL_model_Simple <- Bayesian_Survival_model(
  stan_data = Data_PL_Simple,
  baseline_modelling = "none",
  shrinkage = TRUE,
  withPrediction = FALSE,
  havingInt = TRUE,
  niter = 10000,
  nwarmup = 2000,
  thin = 10,
  chains = 1
)

PL_results_Simple = Bayesian_Survival_result_Extract(
  bayesian_model_fit = PL_model_Simple,
  model_type = "none",
  criteria = c("DesignCoefficients",
               "variableSelection"),
  uncertainty = TRUE
)
PL_Simple_beta <- PL_results_Simple$Beta_bayesian_est[, "mean"]
PL_Simple_variableSelection <- PL_results_Simple$variableSelection
PL_Simple_coefUB <- PL_results_Simple$Beta_bayesian_estUB
PL_Simple_coefLB <- PL_results_Simple$Beta_bayesian_estLB


#---------No shrinkage----------------
bSplines_model_noshrinkage_Simple <-
  Bayesian_Survival_model(
    stan_data = Data_bSplines_Simple,
    baseline_modelling = "bSplines",
    shrinkage = FALSE,
    withPrediction = FALSE,
    havingInt = TRUE,
    niter = 10000,
    nwarmup = 2000,
    thin = 10,
    chains = 1
  )
    
# Extract the model result
bSplines_model_noshrinkage_results_Simple = Bayesian_Survival_result_Extract(
  bayesian_model_fit = bSplines_model_noshrinkage_Simple,
  model_type = "bSplines",
  criteria = c(
    "DesignCoefficients",
    "variableSelection",
    "baseline"
  ),
  uncertainty = TRUE
)
    
bSplines_model_noshrinkage_Simple_beta <- bSplines_model_noshrinkage_results_Simple$Beta_bayesian_est[, "mean"]
bSplines_model_noshrinkage_Simple_variableSelection <- bSplines_model_noshrinkage_results_Simple$variableSelection
bSplines_model_noshrinkage_Simple_coefUB <- bSplines_model_noshrinkage_results_Simple$Beta_bayesian_estUB
bSplines_model_noshrinkage_Simple_coefLB <- bSplines_model_noshrinkage_results_Simple$Beta_bayesian_estLB

PL_model_noshrinkage_Simple <- Bayesian_Survival_model(
  stan_data = Data_PL_Simple,
  baseline_modelling = "none",
  shrinkage = FALSE,
  withPrediction = FALSE,
  havingInt = TRUE,
  niter = 10000,
  nwarmup = 2000,
  thin = 10,
  chains = 1
)

PL_results_noshrinkage_Simple = Bayesian_Survival_result_Extract(
  bayesian_model_fit = PL_model_noshrinkage_Simple,
  model_type = "none",
  criteria = c("DesignCoefficients",
               "variableSelection"),
  uncertainty = TRUE
)
PL_Simple_noshrinkage_beta <- PL_results_noshrinkage_Simple$Beta_bayesian_est[, "mean"]
PL_Simple_noshrinkage_variableSelection <- PL_results_noshrinkage_Simple$variableSelection
PL_Simple_noshrinkage_coefUB <- PL_results_noshrinkage_Simple$Beta_bayesian_estUB
PL_Simple_noshrinkage_coefLB <- PL_results_noshrinkage_Simple$Beta_bayesian_estLB

save.image(file = "./Data/04_OverallKidneyData_Final_All.RData")
```

# Assumptation checking
```{r, echo=TRUE}
load(file = "./Data/04_OverallKidneyData_Final_All.RData")
beta = as.vector(bSplines_model_beta); X = as.matrix(combined_Data_int[, !(names(combined_Data_int) %in% c("obstime", "status"))]); times = combined_Data_int$obstime; status = combined_Data_int$status

Schoenfeld_residuals <- Schoenfeld_resid_tied(beta, X, times, status)
```

#Visualize the result
```{r, echo=TRUE}
load(file = "./Data/04_OverallKidneyData_Final_All.RData")
name_label <- names(combined_Data_int)[1:136]
name_label_simple <- names(combined_Data_Simple_int)

par(mfrow = c(1, 2), mar = c(0, 0, 0, 0), oma = c(3, 24, 3, 1))
# Set up the plot without drawing points initially
plot(
  bSplines_model_beta, 1:136, type = "n",
  col = "blue", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(bSplines_model_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightblue", border = NA)
}

# Add beta coefficient points
points(bSplines_model_beta, 1:136, col = "blue", pch = 16)

# Add error bars for beta coefficients
segments(
  bSplines_model_coefLB, 1:136,
  bSplines_model_coefUB, 1:136,
  col = "blue"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
axis(2, at = 1:136, labels = name_label, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesbSplines)")

plot(
  bSplines_model_noshrinkage_beta, 1:136, type = "n",
  col = "darkgreen", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(bSplines_model_noshrinkage_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightgreen", border = NA)
}

# Add beta coefficient points
points(bSplines_model_noshrinkage_beta, 1:136, col = "darkgreen", pch = 16)

# Add error bars for beta coefficients
segments(
  bSplines_model_noshrinkage_coefLB, 1:136,
  bSplines_model_noshrinkage_coefUB, 1:136,
  col = "darkgreen"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
#axis(2, at = 1:136, labels = name_label, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesPL_noinfor)")





par(mfrow = c(1, 2), mar = c(0, 0, 0, 0), oma = c(3, 24, 3, 1))
# Set up the plot without drawing points initially
plot(
  PL_beta, 1:136, type = "n",
  col = "blue", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(PL_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightblue", border = NA)
}

# Add beta coefficient points
points(PL_beta, 1:136, col = "blue", pch = 16)

# Add error bars for beta coefficients
segments(
  PL_coefLB, 1:136,
  PL_coefUB, 1:136,
  col = "blue"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
axis(2, at = 1:136, labels = name_label, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesPL)")

plot(
  PL_noshrinkage_beta, 1:136, type = "n",
  col = "darkgreen", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(PL_noshrinkage_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightgreen", border = NA)
}

# Add beta coefficient points
points(PL_noshrinkage_beta, 1:136, col = "darkgreen", pch = 16)

# Add error bars for beta coefficients
segments(
  PL_noshrinkage_coefLB, 1:136,
  PL_noshrinkage_coefUB, 1:136,
  col = "darkgreen"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
#axis(2, at = 1:136, labels = name_label, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesPL_noinfor)")
```



```{r, echo=TRUE}
name_label_simple <- names(combined_Data_Simple_int)[1:66]

par(mfrow = c(1, 2), mar = c(0, 0, 0, 0), oma = c(3, 24, 3, 1))
# Set up the plot without drawing points initially
plot(
  bSplines_model_Simple_beta, 1:66, type = "n",
  col = "blue", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(bSplines_model_Simple_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightblue", border = NA)
}

# Add beta coefficient points
points(bSplines_model_Simple_beta, 1:66, col = "blue", pch = 16)

# Add error bars for beta coefficients
segments(
  bSplines_model_Simple_coefLB, 1:66,
  bSplines_model_Simple_coefUB, 1:66,
  col = "blue"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
axis(2, at = 1:66, labels = name_label_simple, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesbSplines) for simpleData")

plot(
  bSplines_model_noshrinkage_Simple_beta, 1:66, type = "n",
  col = "darkgreen", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(bSplines_model_noshrinkage_Simple_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightgreen", border = NA)
}

# Add beta coefficient points
points(bSplines_model_noshrinkage_Simple_beta, 1:66, col = "darkgreen", pch = 16)

# Add error bars for beta coefficients
segments(
  bSplines_model_noshrinkage_Simple_coefLB, 1:66,
  bSplines_model_noshrinkage_Simple_coefUB, 1:66,
  col = "darkgreen"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
#axis(2, at = 1:136, labels = name_label, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesBSpline_noinfor) for simpleData")





par(mfrow = c(1, 2), mar = c(0, 0, 0, 0), oma = c(3, 24, 3, 1))
# Set up the plot without drawing points initially
plot(
  PL_Simple_beta, 1:66, type = "n",
  col = "blue", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(PL_Simple_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightblue", border = NA)
}

# Add beta coefficient points
points(PL_Simple_beta, 1:66, col = "blue", pch = 16)

# Add error bars for beta coefficients
segments(
  PL_Simple_coefLB, 1:66,
  PL_Simple_coefUB, 1:66,
  col = "blue"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
axis(2, at = 1:66, labels = name_label_simple, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesPL)")

plot(
  PL_Simple_noshrinkage_beta, 1:66, type = "n",
  col = "darkgreen", pch = 16,
  xlab = "Estimated Beta", ylab = "",
  main = "", yaxt = "n"
)

# Get the x-axis limits to use for full-width background rectangles
x_limits <- par("usr")[1:2]

# Highlight selected variables with full-width light blue rectangles
for (i in which(PL_Simple_noshrinkage_variableSelection)) {
  rect(x_limits[1], i - 0.5, x_limits[2], i + 0.5, col = "lightgreen", border = NA)
}

# Add beta coefficient points
points(PL_Simple_noshrinkage_beta, 1:66, col = "darkgreen", pch = 16)

# Add error bars for beta coefficients
segments(
  PL_Simple_noshrinkage_coefLB, 1:66,
  PL_Simple_noshrinkage_coefUB, 1:66,
  col = "darkgreen"
)

# Add a vertical reference line at 0
abline(v = 0, col = "red", lty = 2)

# Add vertical y-axis labels
#axis(2, at = 1:136, labels = name_label, las = 2, cex.axis = 1.0)
title(main = "Beta Estimation (BayesPL_noinfor)")
```

```{r, echo=TRUE}
# Step 4. separate the dataset randomly into two and compare the stability of the model across different sub dataset
# Separate the dataset into two small datasets to see the stablility of the estimation of the models
set.seed(123456)

# Number of folds
k <- 10

# Assuming p is the number of coefficients
p <- 105
p1 <- 45

# Preallocate lists to store results across repetitions
bSplines_model_C_INDEX_Simple = bSplines_model_C_INDEX = rep(NA, k)
coxRegression_C_INDEX_Simple = coxRegression_C_INDEX = rep(NA, k)
PL_C_INDEX_Simple = PL_C_INDEX = rep(NA, k)

bSplines_model_beta <- matrix(NA, nrow = k, ncol = p)
bSplines_model_variableSelection <- matrix(NA, nrow = k, ncol = p)
PL_beta <- matrix(NA, nrow = k, ncol = p)
PL_variableSelection <- matrix(NA, nrow = k, ncol = p)
cox_beta <- matrix(NA, nrow = k, ncol = p)
cox_variableSelection <- matrix(NA, nrow = k, ncol = p)
  
  
bSplines_model_beta_Simple <- matrix(NA, nrow = k, ncol = p1)
bSplines_model_variableSelection_Simple <- matrix(NA, nrow = k, ncol = p1)
PL_beta_Simple <- matrix(NA, nrow = k, ncol = p1)
PL_variableSelection_Simple <- matrix(NA, nrow = k, ncol = p1)
cox_beta_Simple <- matrix(NA, nrow = k, ncol = p1)
cox_variableSelection_Simple <- matrix(NA, nrow = k, ncol = p1)  
 

folds <- createFolds(combined_Data_int$status, k = k, list = TRUE)

for (i in seq_along(folds)) {
    indices <- folds[[i]]
    training_dataset <- combined_Data_int[-indices,]
    testing_dataset <- combined_Data_int[indices,]
    
    training_dataset_Simple <- combined_Data_Simple_int[-indices,]
    testing_dataset_Simple <- combined_Data_Simple_int[indices,]
    
    cat("Fold", i, "\n")
    cat("Dataset size:", nrow(training_dataset), "\n")
    
    # insert your model training and evaluation code here
    #-----------MODEL 1.1: bSpline model----------
    Data_bSplines <- stan_data_Constructer_BSpline(
      training_dataset = training_dataset,
      testing_dataset = testing_dataset,
      obs_window = 3415,
      log_lik_saving = FALSE,
      havingInt = TRUE
    )
    bSplines_model <-
      Bayesian_Survival_model(
        stan_data = Data_bSplines,
        baseline_modelling = "bSplines",
        shrinkage = TRUE,
        withPrediction = TRUE,
        havingInt = TRUE,
        niter = 10000,
        nwarmup = 2000,
        thin = 10,
        chains = 1
      )
    
    # Extract the model result
    bSplines_model_results = Bayesian_Survival_result_Extract(
      bayesian_model_fit = bSplines_model,
      model_type = "bSplines",
      criteria = c(
        "DesignCoefficients",
        "variableSelection",
        "Prediction_SurvivalProb",
        "baseline"
      )
    )
    
    bSplines_model_beta[i, ] <- bSplines_model_results$Beta_bayesian_est[, "mean"]
    bSplines_model_variableSelection[i, ] <- bSplines_model_results$variableSelection
    bSplines_model_Prediction_SurvivalProb <- as.vector(bSplines_model_results$eta_pred)
    bSplines_model_C_INDEX[i] <- rcorr.cens(-bSplines_model_Prediction_SurvivalProb, Surv(testing_dataset$obstime, testing_dataset$status))["C Index"]
    baseline_bSplines <-
      exp(as.matrix(Data_bSplines$basis) %*% as.matrix(bSplines_model_results$baselinePara[, "mean"]))
    
    
    
    #-----------MODEL 1.2: Cox regression ----------
    cox_regression <- coxph(Surv(as.numeric(obstime), status) ~ ., data = training_dataset)
    cox_regression_result <- summary(cox_regression)
   
    cox_beta[i,] <- as.vector(cox_regression_result$coefficients[, "coef"])
    cox_variableSelection[i,] <- as.vector((cox_regression_result$coefficients[, "Pr(>|z|)"] * p) < 0.05)
    coxRegression_SurvivalProb <- predict(cox_regression, newdata = testing_dataset, type = "lp")
    coxRegression_C_INDEX[i] <- rcorr.cens(-coxRegression_SurvivalProb, Surv(testing_dataset$obstime, testing_dataset$status))["C Index"] 
    
    #-----------MODEL 1.3: PL model----------
    Data_PL <- stan_data_Constructer_PL(
      training_dataset = training_dataset,
      testing_dataset = testing_dataset,
      havingInt = TRUE
    )
    
    PL_model <- Bayesian_Survival_model(
      stan_data = Data_PL,
      baseline_modelling = "none",
      shrinkage = TRUE,
      withPrediction = TRUE,
      havingInt = TRUE,
      niter = 10000,
      nwarmup = 2000,
      thin = 10,
      chains = 1
    )
    
    PL_results = Bayesian_Survival_result_Extract(
    bayesian_model_fit = PL_model,
    model_type = "none",
    criteria = c(
      "DesignCoefficients",
      "variableSelection",
      "Prediction_SurvivalProb"
    )
  )
    PL_beta[i, ] <- PL_results$Beta_bayesian_est[, "mean"]
    PL_variableSelection[i, ] <- PL_results$variableSelection
    PL_Prediction_SurvivalProb <- PL_results$eta_pred
    PL_C_INDEX[i] <- rcorr.cens(-PL_Prediction_SurvivalProb, Surv(testing_dataset$obstime, testing_dataset$status))["C Index"] 
    
    
    
    
    ###---------------------------------------------------------
    ###-------------for simplified dataset----------------------
    ###---------------------------------------------------------
    #-----------MODEL 2.1: bSpline model----------
    Data_bSplines_Simple <- stan_data_Constructer_BSpline(
      training_dataset = training_dataset_Simple,
      testing_dataset = testing_dataset_Simple,
      obs_window = 3415,
      log_lik_saving = FALSE,
      havingInt = TRUE
    )
    bSplines_model_Simple <-
      Bayesian_Survival_model(
        stan_data = Data_bSplines_Simple,
        baseline_modelling = "bSplines",
        shrinkage = TRUE,
        withPrediction = TRUE,
        havingInt = TRUE,
        niter = 10000,
        nwarmup = 2000,
        thin = 10,
        chains = 1
      )
    
    # Extract the model result
    bSplines_model_results_Simple = Bayesian_Survival_result_Extract(
      bayesian_model_fit = bSplines_model_Simple,
      model_type = "bSplines",
      criteria = c(
        "DesignCoefficients",
        "variableSelection",
        "Prediction_SurvivalProb"
      )
    )
    
    bSplines_model_beta_Simple[i, ] <- bSplines_model_results_Simple$Beta_bayesian_est[, "mean"]
    bSplines_model_variableSelection_Simple[i, ] <- bSplines_model_results_Simple$variableSelection
    bSplines_model_Prediction_SurvivalProb_Simple <- bSplines_model_results_Simple$eta_pred
    bSplines_model_C_INDEX_Simple[i] <- rcorr.cens(-bSplines_model_Prediction_SurvivalProb_Simple, Surv(testing_dataset_Simple$obstime, testing_dataset_Simple$status))["C Index"] 
    
    
    
    #-----------MODEL 1.2: Cox regression ----------
    cox_regression_Simple <- coxph(Surv(as.numeric(obstime), status) ~ ., data = training_dataset_Simple)
    cox_regression_result_Simple <- summary(cox_regression_Simple)
    
    cox_beta_Simple[i,] <- as.vector(cox_regression_result_Simple$coefficients[, "coef"])
    cox_variableSelection_Simple[i,] <- as.vector((cox_regression_result_Simple$coefficients[, "Pr(>|z|)"] * p1) < 0.05)
    coxRegression_SurvivalProb_Simple <- predict(cox_regression_Simple, newdata = testing_dataset_Simple, type = "lp")
    coxRegression_C_INDEX_Simple[i] <- rcorr.cens(-coxRegression_SurvivalProb_Simple, Surv(testing_dataset_Simple$obstime, testing_dataset_Simple$status))["C Index"] 
    
    #-----------MODEL 1.3: PL model----------
    Data_PL_Simple <- stan_data_Constructer_PL(
      training_dataset = training_dataset_Simple,
      testing_dataset = testing_dataset_Simple,
      havingInt = TRUE
    )
    
    PL_model_Simple <- Bayesian_Survival_model(
      stan_data = Data_PL_Simple,
      baseline_modelling = "none",
      shrinkage = TRUE,
      withPrediction = TRUE,
      havingInt = TRUE,
      niter = 10000,
      nwarmup = 2000,
      thin = 10,
      chains = 1
    )
    
    PL_results_Simple = Bayesian_Survival_result_Extract(
      bayesian_model_fit = PL_model_Simple,
      model_type = "none",
      criteria = c(
        "DesignCoefficients",
        "variableSelection",
        "Prediction_SurvivalProb"
      )
    )
    PL_beta_Simple[i, ] <- PL_results_Simple$Beta_bayesian_est[, "mean"]
    PL_variableSelection_Simple[i, ] <- PL_results_Simple$variableSelection
    PL_Prediction_SurvivalProb_Simple <- PL_results_Simple$eta_pred
    PL_C_INDEX_Simple[i] <-
      rcorr.cens(
        -PL_Prediction_SurvivalProb_Simple,
        Surv(testing_dataset_Simple$obstime, testing_dataset_Simple$status)
      )["C Index"] 

}

save.image(file = "./Data/03_results_combined_Step5_includSimplifiedDatanprediction.RData")
```


```{r, echo=TRUE}
load(file = "./Data/03_results_combined_Step5_includSimplifiedDatanprediction.RData")
par(mfrow = c(1, 2))

C_index <- list(bSplines = bSplines_model_C_INDEX, CoxRegression = coxRegression_C_INDEX,  PL = PL_C_INDEX)
boxplot(C_index, names = c("BayesianBSplines", "CoxRegression", "BayesianPL"), col = c("orange", "green","purple"), main = "C-index for complex dataset", xlab = "", ylab = "C-index", ylim = c(0, 1))
abline(h = 0.5, col = "grey", lty = 3)

C_index_Simplified <- list(bSplines = bSplines_model_C_INDEX_Simple, CoxRegression = coxRegression_C_INDEX_Simple, PL = PL_C_INDEX_Simple)
boxplot(C_index_Simplified, names = c("BayesianBSplines", "CoxRegression", "BayesianPL"), col = c("orange", "green","purple"), main = "C-index for complex dataset", xlab = "", ylab = "C-index", ylim = c(0, 1))
abline(h = 0.5, col = "grey", lty = 3)
```


```{r, echo=TRUE}
# Visualize the model results
par(mfrow = c(3, 2), mar = c(2, 2, 2, 1), oma = c(0, 0, 0, 0))
bSplines_model_beta_df <- as.data.frame(t(bSplines_model_beta))
names(bSplines_model_beta_df) <- c("-fold 1", "-fold 2", "-fold 3", "-fold 4", "-fold 5", "-fold 6", "-fold 7", "-fold 8", "-fold 9", "-fold 10")
bSplines_model_beta_cofficients <- cor(bSplines_model_beta_df)
corrplot(bSplines_model_beta_cofficients, method = "color", tl.cex = 1.5, type = "lower", addCoef.col = "white", number.cex = 0.8)
# Add a title to the plot
title("Correlation Plot of BayesianbSpline estimated betas (complex dataset)")


bSplines_model_beta_Simple_df <- as.data.frame(t(bSplines_model_beta_Simple))
names(bSplines_model_beta_Simple_df) <- c("-fold 1", "-fold 2", "-fold 3", "-fold 4", "-fold 5", "-fold 6", "-fold 7", "-fold 8", "-fold 9", "-fold 10")
bSplines_model_beta_Simple_cofficients <- cor(bSplines_model_beta_Simple_df)
corrplot(cor(bSplines_model_beta_Simple_cofficients), method = "color", tl.cex = 1.5, type = "lower", addCoef.col = "white", number.cex = 0.8)
# Add a title to the plot
title("Correlation Plot of BayesianbSpline estimated betas (simplified dataset)")


cox_beta_df <- as.data.frame(t(cox_beta))
names(cox_beta_df) <- c("-fold 1", "-fold 2", "-fold 3", "-fold 4", "-fold 5", "-fold 6", "-fold 7", "-fold 8", "-fold 9", "-fold 10")
cox_beta_cofficients <- cor(cox_beta_df)
corrplot(cox_beta_cofficients, method = "color", tl.cex = 1.5, type = "lower", addCoef.col = "white", number.cex = 0.8)
title("Correlation Plot of cox regression estimated betas (complex dataset)")


cox_beta_Simple_df <- as.data.frame(t(cox_beta_Simple))
names(cox_beta_Simple_df) <- c("-fold 1", "-fold 2", "-fold 3", "-fold 4", "-fold 5", "-fold 6", "-fold 7", "-fold 8", "-fold 9", "-fold 10")
cox_beta_Simple_cofficients <- cor(cox_beta_Simple_df)
corrplot(cox_beta_Simple_cofficients, method = "color", tl.cex = 1.5, type = "lower", addCoef.col = "white", number.cex = 0.8)
title("Correlation Plot of cox regression estimated betas (simplified dataset)")

PL_beta_df <- as.data.frame(t(PL_beta))
names(PL_beta_df) <- c("-fold 1", "-fold 2", "-fold 3", "-fold 4", "-fold 5", "-fold 6", "-fold 7", "-fold 8", "-fold 9", "-fold 10")
PL_beta_cofficients <- cor(PL_beta_df)
corrplot(PL_beta_cofficients, method = "color", tl.cex = 1.5, type = "lower", addCoef.col = "white", number.cex = 0.8)
title("Correlation Plot of Bayesian PL estimated betas (complex dataset)")


PL_beta_Simple_df <- as.data.frame(t(PL_beta_Simple))
names(PL_beta_Simple_df) <- c("-fold 1", "-fold 2", "-fold 3", "-fold 4", "-fold 5", "-fold 6", "-fold 7", "-fold 8", "-fold 9", "-fold 10")
PL_beta_Simple_cofficients <- cor(PL_beta_Simple_df)
corrplot(PL_beta_Simple_cofficients, method = "color", tl.cex = 1.5, type = "lower", addCoef.col = "white", number.cex = 0.8)
title("Correlation Plot of cox regression estimated betas (simplified dataset)")
```

```{r, echo=TRUE}
bSplines_model_beta_corVal <- bSplines_model_beta_cofficients[lower.tri(bSplines_model_beta_cofficients)]
bSplines_model_beta_Simple_corVal <- bSplines_model_beta_Simple_cofficients[lower.tri(bSplines_model_beta_Simple_cofficients)]

cox_beta_corVal <- cox_beta_cofficients[lower.tri(cox_beta_cofficients)]
cox_beta_Simple_corVal <- cox_beta_Simple_cofficients[lower.tri(cox_beta_Simple_cofficients)]

PL_beta_corVal <- PL_beta_cofficients[lower.tri(PL_beta_cofficients)] 
PL_beta_Simple_corVal <- PL_beta_Simple_cofficients[lower.tri(PL_beta_Simple_cofficients)]

beta <- list(bSplines = bSplines_model_beta_corVal, CoxRegression = cox_beta_corVal, PL = PL_beta_corVal)
beta_Simplified <- list(bSplines = bSplines_model_beta_Simple_corVal, CoxRegression = cox_beta_Simple_corVal, PL = PL_beta_Simple_corVal)

par(mfrow = c(1, 2))
boxplot(beta, names = c("BayesianBSplines", "CoxRegression", "BayesianPL"), col = c("orange", "green","purple"), main = "correlation of estimated betas  for complex dataset", xlab = "", ylab = "C-index", ylim = c(0, 1))

boxplot(beta_Simplified, names = c("BayesianBSplines", "CoxRegression", "BayesianPL"), col = c("orange", "green","purple"), main = "C-index for simplified dataset", xlab = "", ylab = "C-index", ylim = c(0, 1))
```


```{r, echo=TRUE}
load("./Data/04_OverallKidneyData_Final_All.RData")
baselinehazard <- exp(as.matrix(Data_bSplines$basis) %*% bSplines_model_results$baselinePara[[1]])
baselinehazard_Simple <- exp(as.matrix(Data_bSplines_Simple$basis) %*% bSplines_model_results_Simple$baselinePara[[1]])
baselinehazard_no <- exp(as.matrix(Data_bSplines$basis) %*% bSplines_model_noshrinkage_results$baselinePara[[1]])
baselinehazard_Simple_no <- exp(as.matrix(Data_bSplines_Simple$basis) %*% bSplines_model_noshrinkage_results_Simple$baselinePara[[1]])

par(mfrow = c(2, 2), mar = c(2, 2, 2, 1), oma = c(0, 0, 0, 0))
timevec <- sort(unique(combined_Data_int$obstime))
plot(timevec, baselinehazard, col = "red", pch = 16, lty = 1, lwd = 2, xlab = "time", ylab = "")
legend(
    "topright",
    legend = "B-Splines",
    col = "red",
    pch = 16,
    lty = 1)
plot(timevec, baselinehazard_no, col = "blue", pch = 16, lty = 1, lwd = 2, xlab = "time", ylab = "")
legend(
    "topright",
    legend = "B-Splines_No",
    col = "blue",
    pch = 16,
    lty = 1)

plot(timevec, baselinehazard_Simple, col = "green", pch = 16, lty = 1, lwd = 2)
legend(
    "topright",
    legend = "B-Splines_Simple",
    col = "green",
    pch = 16,
    lty = 1)

plot(timevec, baselinehazard_Simple_no, col = "orange", pch = 16, lty = 1, lwd = 2)
# Add legend
legend(
    "topright",
    legend = "B-Splines_No_Simple",
    col = "orange",
    pch = 16,
    lty = 1)


```